---
title: "Homework 1"
author: "Kai Aragaki"
output:
  tufte::tufte_handout: default
---

# Problem 1
   a. iii - the mean and median are about the same. The distribution is roughly
      symmetric, and there do not appear to be any large outliers (which the mean is sensitive to).
   b. Were this a normal distribution, we could assume that a standard deviation
      would encapsulate roughly 2/3s of the data present. By eye, a range that
      surrounds the putative mean (~100) by +/- 25 appears to capture roughly
      2/3s of these data, while +/- 10 appears to capture too little and +/- 50
      too much.

# Problem 2

We will assume that when our doctor talked about infection or repair failure risk, he was talking about the risk of getting ONLY an infection or ONLY a fail to repair, not both at the same time. If that is the case:

```{r chances}
data.frame(type = c("infection", "failure", "both", "neither"),
           risk = c("3%", "14%", "1%", "X"))
```
So 

```{r chance_calc}
x <- 100 - 3 - 14 - 1
x
```

`r x`% of these opporations are successful and infection-free.

# Problem 3

Suppose there is a group of 1,000,000 people, and one (and only one) of these people has the variant. With a specificity of 99.99%, that means that

```{r num_pos}
false_pos <- 1000000 * 0.0001
false_pos
```

`r false_pos` will receive results telling them they are positive for this variant, even though they are not! This means that a positive test result means relatively little, as most of the time a positive result will simply be a false positive. 

# Problem 4


